SunPy - Python for Solar Physics

### 2014-01-30 1-Introduction.tex

Science is driven by the analysis of data of growing variety and complexity.
Modern advances in sensor
technology, combined with the availability of inexpensive storage, has led to
rapid increases in the amount of data available to scientists in almost
every discipline.  Solar physics is no exception to this trend. For example,
NASA's \textit{Solar Dynamics Observatory} (\textit{SDO}) spacecraft, launched in February 2010,
produces over 1 TB of data per day \citep{aia}. Managing and
analysing this mountain of data
requires increasingly sophisticated software tools.
These tools should be robust, easy of use and modify, have a
transparent development history, and conform
to modern software engineering standards.  Software with these qualities provide a strong
foundation and can support the needs of the community as data
volumes grow and science questions evolve.

>>>>

Science is driven by the analysis of data of growing variety and complexity.
Modern advances in sensor
technology, combined with the availability of inexpensive storage, has led to
rapid increases in the amount of data available to scientists in almost
every discipline.  Solar physics is no exception to this trend. For example,
NASA's \textit{Solar Dynamics Observatory} (\textit{SDO}) spacecraft, launched 
in February 2010, produces over 1 TB of data per day \citep{aia}. Managing and
analysing this mountain of data requires increasingly sophisticated software 
tools.
These tools should be robust, easy of use and modify, have a transparent 
development history, as well as conforming to modern software engineering 
standards.
Software with these qualities provide a strong foundation which can support the 
needs of the community as data volumes grow and science questions evolve.

### 2014-01-30 3-3-HELIO.tex

The HELiophysics Integrated Observatory (HELIO\footnote{For more information see \url{http://helio-vo.eu}{http://helio-vo.eu}} has 
compiled a list of web services which allows scientists to query and 
discover data throughout the heliosphere, from solar and magnetospheric data to planetary and 
inter-planetary data \citep{dps2012}.
HELIO is built with a Service-Oriented Architecture, 
i.e., its capabilities are divided into a number of tasks that are 
implemented as separate services. 
HELIO is made up of nine different public services, 
which allows scientists to search different catalogues of registered events, 
solar features, data from instruments in the heliosphere and other information 
such as planetary or spacecraft position in time. 
Additionally, HELIO provides a service that uses a 
propagation model to link the data in different points of the solar system by 
its original nature (e.g., Earth auroras are a signature of magnetic 
field disturbances produced few days before on the Sun).
In addition to the primary, web-based interface to 
HELIO, the services are available by API.

>>>>

The HELiophysics Integrated Observatory (HELIO\footnote{For more information 
see \url{http://helio-vo.eu}{http://helio-vo.eu}}) has 
compiled a list of web services which allows scientists to query and 
discover data throughout the heliosphere, from solar and magnetospheric data to planetary and 
inter-planetary data \citep{dps2012}.
HELIO is built with a Service-Oriented Architecture, 
i.e., its capabilities are divided into a number of tasks that are 
implemented as separate services. 
HELIO is made up of nine different public services, 
which allows scientists to search different catalogues of registered events, 
solar features, data from instruments in the heliosphere and other information 
such as planetary or spacecraft position in time. 
Additionally, HELIO provides a service that uses a 
propagation model to link the data in different points of the solar system by 
its original nature (e.g., Earth auroras are a signature of magnetic 
field disturbances produced few days before on the Sun).
In addition to the primary, web-based interface to 
HELIO, the services are available by API.

### 2014-01-30 3-Retrieval.tex

Several well-developed resources currently exist within heliophysics which provide  
remote access and data retrieval for a large number of data sources as well as access to 
event databases. As a community-based project, 
SunPy provides support for these resources via the \texttt{net} subpackage. 
In the following subsections, we describe a number of these resources and 
how to use them in SunPy.

>>>>

Several well-developed resources currently exist which provide remote access to 
and data retrieval form a large number of solar and heliospheric data sources 
and event databases. SunPy provides support for these resources via the 
\texttt{net} subpackage. In the following subsections, we describe a number of 
these resources and how to use them in SunPy.

### 2014-01-30 abstract.tex

\begin{abstract}
This paper presents version 0.4 of SunPy a community developed Python package 
for solar physics.

\end{abstract}

>>>>

\begin{abstract}
This paper presents version 0.4 of SunPy, a community-developed Python package 
for solar physics and data analysis.
Python, a free, cross platform, general purpose, high-level programming 
language, has seen widespread adoption among the scientific community resulting 
in the availability of a large range of available software, from numerical 
computation (NumPy, SciPy) and machine learning (scikit-learn) to spectral 
analysis and visualisation (Matplotlib).
SunPy is a data analysis toolkit specialising in providing the software 
necessary to analyse solar and heliospheric datasets in Python. 
SunPy is open source software (BSD licence) and has a open and transparent 
development workflow that anyone can contribute to.
SunPy provides access to solar data through integration with the Virtual 
Solar Observatory (VSO), the the Heliophysics Event Knowledgebase (HEK) and the 
HELIO webservices. It can open data files from major solar missions (e.g. 
SDO, SOHO, STEREO and IRIS) into WCS-aware `Maps'. Time-series data from 
mission such as GOES, SDO/EVE, and Proba2/LYRA into `LightCurves' and radio 
spectra from e-Callisto or STEREO/SWAVES into `Spectrograms'. We present 
examples of solar data analysis in SunPy, and show how Python-based solar 
data-analysis can leverage the many existing data analysis tools already 
available in Python. We discuss the future goals of the project and encourage 
interested users to become involved in the planning and development of SunPy.
\end{abstract}
